---
title: "Clustering"
author: "YOUR NAME HERE"
format: 
  html:
    code-fold: true
    code-line-numbers: true
    code-tools: true
    self-contained: true
editor: visual
execute:
  message: false
---

## Setup

Declare your libraries:

```{r}
#| label: libraries-r
#| include: false
library(tidyverse)
library(tidymodels)
library(glmnet)
library(discrim)
library(rpart)
library(rpart.plot)
library(baguette)
```


## Data Prep

```{r}
cann <- read_csv("https://www.dropbox.com/s/s2a1uoiegitupjc/cannabis_full.csv?dl=1")

cann <- cann %>% drop_na()

cann_matrix <- cann %>%
  select(-Type, -Strain, -Effects, -Flavor, -Dry, -Mouth) %>%
  as.matrix()

cann_types <- cann %>% pull(Type)
```


## Code from Lecture

```{r}
fed_km <- kmeans(fed_matrix, 3)
fed_km

fed_km$centers

fed_km$totss
fed_km$withinss
fed_km$betweenss

res <- tibble(
  clust = fed_km$cluster, 
  auth = auths_known)

res

res %>% count(clust, auth)
```


```{r}
pc <- prcomp(fed_matrix, center = TRUE, scale = TRUE)

fed_reduced <- pc$x[, 1:2]

fed_pca_km <- kmeans(fed_reduced, 3)

res <- tibble(
  clust = fed_pca_km$cluster, 
  auth = auths_known)

res %>% count(clust, auth)
```

```{r}

fed_pca_km <- kmeans(fed_reduced, 4)

res <- tibble(
  clust = fed_pca_km$cluster, 
  auth = auths_known)

res %>% count(clust, auth)
```



## Try it!

#### Apply k-means clustering to the cannabis data using *all* the word predictors.
#### What was the within and between sum of squares?
#### Did the clusters match up with the Type?

#### Refer back to your PCA analysis of the cannabis data.
#### Apply k-means clustering to the **second and third** PC only
#### Plot these clusters.  What do you think they capture?

## More code from class

```{r}
fed_hc <- fed_matrix %>% 
  scale() %>% 
  dist() %>% 
  hclust()

plot(fed_hc, labels = fed_known$auths)
```



```{r}
res_hc <- cutree(fed_hc, k = 3)

res_hc
```

```{r}
tibble(
  clust = res_hc,
  auth = fed_known$auths
) %>%
  count(clust, auth)
```

```{r}
res_hc_2 <- cutree(fed_hc, h = 0.05)
res_hc_2
```

```{r}
tibble(
  clust = res_hc_2,
  auth = auths_known
) %>%
  count(clust, auth)
```

## Try it!

#### Apply hierarchical clustering to the cannabis data
#### Compare your results to k-means. Which do you prefer?  Why?
